<!-- Instructions: This is the template you will use to type up your responses to the exercises and the on your own questions. To produce a document that you can print out and turn in just click on Knit HTML above. All you need to do to complete the lab is to type up your BRIEF answers and the R code (when necessary) in the spaces provided below. If you want to find out more about the markdown language click on the MD icon. -->

Lab 9
========================================================

#### Author: Colm Kennedy


**Load data:**

```{r}
#install.packages('printr')
#install.packages('xtable')
require('printr')
require('xtable')
#download.file('https://mmmlab.org/courses/psych400/data/mlb11.RData',
#              destfile='~/R/data/mlb11.RData')
load("~/R/data/mlb11.RData")
```


#### Exercise 1:

In order to display the relationship between two numerical values (at_bats and runs), I would use a scatterplot. In the plot below, I observed a mild positive linear relationship. In the absence of more highly correlated variables, I would feel comfortable using a linear model to predict the number of runs. There is a linear correlation visible, so it is likely the model would provide some information about runs. Because the correlation appears not to be incredibly strong, I would prefer to find a variable with a stronger correlation.

<!--Insert any code that should be run in the console into the block below-->
```{r}
# use scatterplot
plot(mlb11$at_bats, mlb11$runs)
cor(mlb11$runs, mlb11$at_bats)
```

#### Exercise 2:

Runs and at_bats have a positive linear correlation. The correlation is 0.611, which is a moderately strong correlation. There is an unusual observation around 5500 at_bats, the matched number of runs is much higher than the rest of the data, at around 850 runs. We would expect somewhere in the region of 700 runs for an at_bats of 5500.

#### Exercise 3:

After running plot_ss several times, the best sum of squares I could obtain was 126024.3. My first try was 129803, then on my second try I got my best score. Each attempt beyond that was much higher, going as high as 187887.

<!--Insert any code that should be run in the console into the block below-->
```{r}
plot_ss(x = mlb11$at_bats, y = mlb11$runs, showSquares = TRUE)

```


#### Exercise 4:

- Estimate dived by standard error is the t value
- larger t value means stronger predictor
- r^2 tells you the same

y_hat = 415.2389 + 1.8345 * homeruns

The slope tells us how many more runs we expect from one additional home run. According to this model, for each home run, we expect a 1.8345 increase in total runs. This makes sense, because a home run guarantees at least one run, and our slope is greater than 1. I am not entirely sure on my knowledge of baseball, but I believe the maximum amount of runs possible from one home run is 4. Our slope is also less than 4. It is logical that a home run should increase total runs by 1.8345, because: 1 < 1.8345 < 4. 

<!--Insert any code that should be run in the console into the block below-->
```{r}
# runs on at_bats
m1 <- lm(runs ~ at_bats, data = mlb11)

# runs on homeruns
m2 <- lm(runs ~ homeruns, data = mlb11)
#summary(m2)
xtable(summary(m2))

# equation of regression line in the form of a function
y_hat_homeruns <- function(homeruns){
  # Takes number of homeruns, outputs expected number of runs
  return(415.2389 + 1.8345 * homeruns)
}

```

#### Exercise 5:

The team manager would expect 729 runs. This is an overestimate of 16 runs, the actual amount of runs for a team with 5579 at bats is 713. This also means the residual for this prediction is 16 runs. 

<!--Insert any code that should be run in the console into the block below-->
```{r}
plot(mlb11$runs ~ mlb11$at_bats)
abline(m1)


prediction = m1$coefficients[2]*5579+m1$coefficients[1]
prediction

mlb11$runs[mlb11$at_bats==5579]

# dont use this for lab, just for checking
#plot(m1)
```

#### Exercise 6:

The observed number of runs (713) for the team with 5579 at-bats falls in the percentile interval. In fact, it falls very close to the regression line.

<!--Insert any code that should be run in the console into the block below-->
```{r}
plot_PI(mlb11$at_bats,mlb11$runs, 0.95)
# adding a line at 5,579 at-bats and 713 runs to aid in visualization
abline(v=5579, col='blue')
abline(h = 713, col = 'blue')
```

#### Exercise 7:

There are no apparent patterns in the variability. The residuals appear to be relatively evenly distributed and there is no clear cone shape or other recognizable pattern in deviation. There is still one outlier around 5525 at bats, which is a much higher residual than the rest. This tells us the relationship between runs and at-bats is linear. The residual has relatively equal variance, which provides evidence of a linear relationship.

<!--Insert any code that should be run in the console into the block below-->
```{r}
plot(m1$residuals ~ mlb11$at_bats)
abline(h = 0, lty = 3)  # adds a horizontal dashed line at y = 0
```

#### Exercise 8:

According to the plots, the approximately normal residuals condition appears  to be met. The histogram appears to have a mean close to 0, and the tails have a lower frequency, following a roughly bell-like shape. The qqplot provides a bit more information on the variance from normal. As expected, there is more variance from normal in the tails, but it is not extreme. There does appear to be somewhat of a stepwise pattern in the variance, where we can see several residual points in a horizonal line around -1, 0, and 1 on the x-axis. Because this stepwise pattern appears systematic, it is a point of concern, but I believe the residuals still follow the normal probability line closely enough to meet the approximately normal condition. 

<!--Insert any code that should be run in the console into the block below-->
```{r}
hist(m1$residuals)

qqnorm(m1$residuals)
qqline(m1$residuals)
```

#### Exercise 9:

Based on the residual plot, constant variance appears to be met. There is no clear shape or pattern in the residual points.

<!--Insert any code that should be run in the console into the block below-->
```{r}

```

#### Exercise 10:

The confidence interval is (0.00014, 1.26086)

<!--Insert any code that should be run in the console into the block below-->
```{r}
# CI = b +/- t * (df) * SE
xtable(summary(m1))

lower <- 0.6305 - 4.08 * 0.1545
upper <- 0.6305 + 4.08 * 0.1545
lower
upper
```
* * *

### On your own:


#### 1:

At a glance, there appears to be a very strong linear relationship between bat_avg and runs. The correlation between batting average and runs is extremely high at: 0.81

<!--Insert any code that should be run in the console into the block below-->
```{r}
# run on new_onbase
plot(mlb11$bat_avg, mlb11$runs)
m3 <- lm(runs~bat_avg, data = mlb11)
cor(mlb11$runs,mlb11$bat_avg)
```

#### 2:

Comparing my model, runs on batting average, I received an r^2 of 0.6561, compared to 0.3729 from the runs on at-bat model. This is much higher, and means my model accounts for 65.61% of the variability in runs. This is a 28.32% increase in variability in runs explained by batting average. Based on this, my model is better at predicted runs than the previous model.

<!--Insert any code that should be run in the console into the block below-->
```{r}
#summary(m1)
#summary(m3)
xtable(summary(m1))
xtable(summary(m3))
```

#### 3:

After investigating the other five variables, my model (m3), runs on batting average, is the best model. With an r^2 of 0.6561, it has a higher r^2 than any other model with the traditional variables. 

<!--Insert any code that should be run in the console into the block below-->
```{r}
names(mlb11)
m4 <- lm(runs~hits, data = mlb11)
m5 <- lm(runs~strikeouts, data = mlb11)
m6 <- lm(runs~stolen_bases, data = mlb11)
m7 <- lm(runs~wins, data = mlb11)

# summary(m6)
# summary(m2)
# summary(m7)
# summary(m4)
# summary(m5)

#summary(m3)
xtable(summary(m3))

plot_PI(mlb11$bat_avg,mlb11$runs, 0.95)

```

#### 4:

new_obs seems to be the best overall predictor of runs. On base slugging on runs has the best r^2 of any model, at 0.9349. This means on base slugging explains 93.49% of variability in runs. This is an increase of 27.88% from the best model I had found in the traditional variables, batting average. We can also see from the plot, the confidence interval is much tighter than the the CI was for batting average. This makes on base slugging the best predictor we have of runs. This makes sense, because OBS is the ability for a player to both get on base, and hit for power. This means the player is more likely to get themselves on base, and get the ball far enough away to allow others to make runs. A player with a high OBS would then be able to not only score themselves, but allow other players to score. 

<!--Insert any code that should be run in the console into the block below-->
```{r}
names(mlb11)
m8 <- lm(runs~new_onbase, data = mlb11)
m9 <- lm(runs~new_slug, data = mlb11)
m10 <- lm(runs~new_obs, data = mlb11)

#summary(m8)
#summary(m9)
xtable(summary(m10))

plot_PI(mlb11$new_obs,mlb11$runs, 0.95)

```

#### 5:

The model for runs on OBS seems homeostatic. The residual plot shows equal variance throughout every interval of obs. There are no clear patterns in the residual. The histogram of the residuals appears approximately normal, the mean appears to be close to 0, and the distribution appears close to bell-shaped. The qq plot also supports a normal distribution. The residuals follow the normal distribution line very closely. The residual plot also supports linearity, showing the residuals are mostly evenly distributed.  

<!--Insert any code that should be run in the console into the block below-->
```{r}
plot(m10$residuals ~ mlb11$new_obs)
abline(h = 0, lty = 3) 

hist(m10$residuals)

qqnorm(m10$residuals)
qqline(m10$residuals)
```



